---
title: "U.S. Fish and Wildlife Imports"
output: github_document
theme: sandstone
fig_caption: yes
df_print: tibble
html_notebook: default
pdf_document: default
---

This dataset describing wildlife imports between 2000-2014 was collected by the U.S. Fish and Wildlife Services and cleaned by Law Enforcement Management Information Systems (LEMIS). The EDA below focused on the fish taxa.

---

Installing packages and loading libraries

```{r}

library(ggplot2)
library(tidyverse)
library(skimr)


```


Set the working directory
```{r}
setwd("~/Documents/R/USFWS")

```

Import Dataset
```{r}
lemis<-read_csv("lemis.csv")
```


Take a quick overview of which taxa has the most observations
```{r}
str(lemis)
summary(lemis)


lemis %>%
  count(taxa)


```
I am choosing fish as a taxa because it appears to have enough data to work with.

Now I will create a new dataset only looking at the fish taxa, looking it over 
before I begin any plotting or manipulation.
```{r}


fish <- lemis %>%
  filter(taxa=="fish")

str(fish)
summary(fish)

```
Now I want to narrow my dataset to only observe value (of import) and import year within the fish taxa. 
```{r}
smfish <-fish %>%
  select(taxa, value, shipment_year)
  
skim(smfish)
summary(smfish)



```

I want to know if there were certain years when fish imports slowed or skyrocketed, basically how has the fish trade varied over time? 

Looking at the number of fish imports per year, we have the following plot.
```{r}
ggplot(smfish, aes(shipment_year)) + geom_bar() + xlab ("Year") + ylab ("Number of Fish Shipments to US") + ggtitle(label="Imported Fish Shipments Between 2000-2014", subtitle="Katy Baker")

```


Now I wanted to look at the range of declared value of the fish shipments to the US over time. 
I wondered if the fish trade is dominated by few, expensive fish products, or many, lower cost fish products. 

I use the jitter to reduce overplotting.

```{r}
ggplot(smfish, aes(taxa, value)) + geom_jitter(height=0, width= .5, alpha=.7, size=1, color="blue")

```


I wanted to see what that outlier was. So I used the filter function to look
at all the observations over 10 million dollars.

```{r}
outlier<-lemis %>%
  filter(value>1e7 & taxa=="fish") %>%
  arrange(value)
outlier
```

Turns out the most extreme value was the import of just one Tiger Shark from 
Australia worth 85 million dollars.
I thought it could be in error so I did a quick google on the seller (forein_co)
and apparently it was a piece of art by Damien Hirst called The Physical 
Impossibility of Death in the Mind of Someone Living (1991).

It was literally a tiger shark preserved in a glass tank of formaldehyde. 
For 85 million dollars. 

Not only that but according to this article,  https://www.thegentlemansjournal.com/5-hedge-fund-billionaire-steve-cohens-extravagant-purchases/
the buyer, a hedge-fund billionaire, purchased it in 2006 and didn't even have it 
shipped to the US until 2012, per LEMIS.

For the purpose of this data exploration we are not going to include this 
ridiculous display of wealth which I am obviously jealous of and which as made me
rethink all of my life choices.

We are going to include the 10 million dollar caviar, though, since it is actually
relevant to this study, albeit ridiculous as well. 

I create a new dataframe and use the filter function to only include observations 
where the value is less than 20M

```{r}
smsmfish<- smfish %>%
  filter(value<2e7) 
  
```


Then I wanted to also get rid of the '0' value, so I will use the filter function
in the same way after I make sure there are no values between 0 and 1

```{r}
outlier2<-lemis %>%
  filter(value<1 & taxa=="fish") %>%
  arrange(desc(value))
outlier2

smsmfish<- smfish %>%
  filter(between(value,1, 2e7))


skim(smsmfish)

```


Then plotting again except with more extreme jitter to reduce overplotting

```{r}
ggplot(smsmfish, aes(taxa, value)) + geom_jitter(height=0, width= .5, alpha=.7, size=1, color="blue") + xlab ("") + ylab ("Value of Fish Shipments") + ggtitle(label="Declared Value of Imported Fish Shipments Between 2000-2014", subtitle="Katy Baker")


smsmfish<- smfish %>%
  filter(between(value,5, 2500))

ggplot(smsmfish, aes(taxa, value)) + geom_boxplot(color="blue") + xlab ("") + ylab ("Value of Fish Shipments") + ggtitle(label="Declared Value of Imported Fish Shipments Under $2500 Between 2000-2014", subtitle="Katy Baker")

```


RESULTS



While the fish trade seems to be relatively stable and did not have much 
volatility, there did appear to be a momentary decline in imported fish from 2006-2010.
Without looking at other variables such as species, it is difficult to begin to 
ask why this might be. For instance, if the decline overall was proportional to 
the decline of imports for a certain species, say tuna, then we might be able to
distinguish whether the downward trend is fish imports overall or just one sector.

The EDA on declared value of fish shipments to the US did, however, prove to be 
informative in its own right. It seems clear that the fish import trade is dominated 
by smaller valued imports. There are very expensive fish/products being imported, 
and although they may make up a significant portion of the overall value of the 
fish trade, the average shipment is only valued at 0.03% of the highest valued
import we considered (the 10 million dollar caviar). It was because of this 
extremely right-skewed data that a histogram was not a good choice for depicting 
the spread of values. 

And who doesn't love exploring data to uncover the ultimate American dream: 
getting so filthy rich off an imaginary construct that you become bored with
spending money and resort to buying obscure pieces of taxidermy someone deemed
art that - no doubt - was a royal pain to ship.
